---
title: "Assignament"
author: "Sathish V"
date: "May 28, 2016"
output: html_document
---

###Assignment: Prediction Assignment Writeup

####Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).


####Data Cleaning
The data for this project come was obteined from http://groupware.les.inf.puc-rio.br/har. Two data set were available a training set and a test set for which 20 individuals without any classification for the class of exercise was available

```{r}
rm(list=ls())
setwd('C:/Study/Practical Machine Learning/Project')
fitTrain<-read.csv("pml-training.csv", header=T, na.strings=c("NA", "#DIV/0!"))
fitTest<-read.csv("pml-testing.csv", header=T, na.string=c("NA", "#DIV/0!"))

```

Training data was partitioned and preprocessed using the code described below. In brief, all variables with at least one "NA" were excluded from the analysis. Variables related to time and user information were excluded for a total of 51 variables and 19622 class measurements. Same variables were mainteined in the test data set (Validation dataset) to be used for predicting the 20 test cases provided.

```{r }
## NA exclusion for all available variables
noNApmlTrain<-fitTrain[, apply(fitTrain, 2, function(x) !any(is.na(x)))] 
dim(noNApmlTrain)

## variables with user information, time and undefined
cleanpmlTrain<-noNApmlTrain[,-c(1:8)]
dim(cleanpmlTrain)
## 20 test cases provided clean info - Validation data set
cleanpmltest<-fitTest[,names(cleanpmlTrain[,-52])]
dim(cleanpmltest)
```

###Data Partitioning and Prediction Process
The cleaned data set split into traing ans test data sets.   Partitioning was performed to obtain a 75% training set and a 25% test set.

```{r }

 #data cleaning
library(caret)
inTrain<-createDataPartition(y=cleanpmlTrain$classe, p=0.75,list=F)
training<-cleanpmlTrain[inTrain,] 
test<-cleanpmlTrain[-inTrain,] 
#Training and test set dimensions
dim(training)
dim(test)
```

###Results and Conclusions
Random forest trees were generated for the training dataset and validated with test data set. 

```{r }
library(caret)
set.seed(13333)
 
rffit<-train(classe~.,data=training, method="rf")

predrf<-predict(rffit, newdata=test)
confusionMatrix(predrf, test$classe)

pred20<-predict(rffit, newdata=cleanpmltest)
# Output for the prediction of the 20 cases provided
pred20
```
A boosting algorithm was also run to confirm and be able to compare predictions. The predictions for the 20 test cases were compared and there is 100% match for both ran algorimths.

```{r }
 
gmbfit<-train(classe~.,data=training, method="gbm" )
gmbfit$finalModel
class(gmbfit)
predgmb<-predict(gmbfit, newdata=test)
confusionMatrix(predgmb, test$classe)
predtrain<-predict(gmbfit, newdata=training)
confusionMatrix(predtrain, training$classe)
predtrain<-predict(gmbfit, newdata=training)
confusionMatrix(predtrain, training$classe)

```
 